---
title: "Case Study - Cyprus Tourism prospects."
author: "Antonios Kokiantonis"
date: "31/12/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 

#  Scenario

   Our Data science team was hired by the multinational company Intertourism corporation. 
   Intertourism, owns many hotels worldwide , mostly in Asia and South America. Company's hotels operate 
   mostly as "all inclusive" and offer to their customers attached activity zones like waterparks, tennis
   courts, scuba diving, etc.

  Their board of directors decide to expand in Europe, and contact our company, Idealdata, for further 
  invetigation. 

### ASK

   Our directors, decided that since Intertourism expertise in mostly summer activities, we
   will research Europian countries with such characterisrics. Cyprus, Greece, Italy and Spain will be
   the four countries to start with.
   We will focus our analysis in two 'axis'. Our team will split their efforts in 
   1) Cost of investment:
   find and analyse data about the costs of buying land, building and maintain hotels and parks in 
   each country and 
   2) The return of investment: find and analyse data about the tourist prospectivess in each country.

   Since I am a Greek-Cypriot, I offered my self to work for Cyprus. I made to my team_mates clear that
   I tottaly respect our customers interests, and my work will be unbiased according to my country, 
   and also maybe take advantages in analysis, by the better knowledge I have about my country. 
   We decided that my part in this team work will be: Cyprus tourism characteristics. 

##    Case Study - Cyprus tourism prospects. 
    
####    Objectives
   
   *  Tourists volume and spending habits.
   *  Industry's tendencies.
   *  Corona's impact.

   It will then be compined with my partner's Study: 'Cost of creating hospitality business in Cyprus'.
   so we create the Cyprus_tourism 
   It will then be compined with other teams' Greece_tourism , Italy_tourism , Spain_tourism
   To finally create the Best_option_tourism
   
###    PREPARE

### Data source
 
   Thankfully Cypriot government, https://www.data.gov.cy/ has data about its tourism.  
 
   Thats great news, since a governmental open source  data to covers licensing and credibility,
   probably also its integrity.

   Licence link: https://creativecommons.org/licenses/by/4.0/ 

   I've downloaded 6 files:

   1. AIRPORT ARRIVALS_16.csv
   2. ARRIVALS BY COUNTRY_4.csv
   3. ARRIVALS_17.csv
   4. EXPENDITURE BY MONTH_22.csv
   5. TOURISM REVENUE_23.csv
   6. TOURISMeurorevised JULY2020 WEBb.xls
   
###   PROCESS

#### It is time to explore and clean our data. 
```{r}
install.packages("readr")
library(readr)
```

#### 1st file - Airport arrivals

```{r}
airarrivals <- read.csv(file.choose())
options(scipen = 100)
install.packages("DataExplorer")
library(DataExplorer)
create_report(airarrivals_year_transposed)
```




It comes with an error:
  Quitting from lines 122-129 (report.rmd) 
  Error in (function (data, with = NULL, by = NULL, by_position = "fill", : 
                     ## Note: All discrete features ignored! Nothing to plot!

```{r}
str(airarrivals_year_transposed)
```

The month variable has probably encoding issues, so January01 is written 'Î™Î±Î½-01, etc.
   I must fix this with code:

```{r}
install.packages("stringr")
library(stringr)
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "Î™Î±Î½-", "January-")
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "Î¦ÎµÎ²-", "-February")
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "ÎœÎ±Ï", "March")
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "Î‘Ï€Ï-", "-April")
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "ÎœÎ±ÏŠ", "May")
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "Î™Î¿Ï…Î½", "June")
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "Î™Î¿Ï…Î»", "July")
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "Î‘Ï…Î³-", "August")
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "Î£ÎµÏ€-", "September")
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "ÎŸÎºÏ„", "October")
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "ÎÎ¿Îµ", "November")
airarrivals$MONTH <- str_replace(airarrivals$MONTH, "Î”ÎµÎº", "December")
```


 Since a general Exploratory is not available at the moment by DataExplorer, I will try some of its 
 functions

```{r}
plot_missing(airarrivals)
```


  No missing values found. That is very good
  
```{r}
plot_histogram(airarrivals)
plot_density(airarrivals)
plot_qq(airarrivals)
duplicated(airarrivals)
```
  
   No duplicates found, very good!
   
```{r}
install.packages("SmartEDA")
library(SmartEDA)
ExpReport(airarrivals, op_file = "smarteda.html")
install.packages("dlookr")
library(dlookr)
diagnose_web_report(airarrivals)
diagnose_paged_report(airarrivals)
```



  There is a warning about one zero value at Paphos_airport. After my check it seems that it happened
  during corona lock_down Mar2020. 
  The report shows that: There are not missing values, all variables are unique, there is one 'zero',
  which is justified, no outliers, no negative values, no duplicates. Our data may be considered clean.
  
  
#### Continue cleaning in Spreadsheets

  Since the data files are small I've realized that working on the spreadsheet environment will be faster.
  
