---
title: "Analysis of world Population"
author: "James Akemu"
date: "2023-12-25"
output: html_document
---

# Analysis of World Population

In this Project, We will take a look at the countries and continents of the world and what the data about their population from 1970-2022 tell us. We will conduct a descriptive analysis of our dataset. Firstly let's import and clean our data.
## Load the Necessary packages.

```{r}
library(tidyverse)
```

## Import the data and clean it

```{r}
world_population <- read_csv("other/world_population.csv", 
     col_types = cols(Rank = col_number(), 
         Continent = col_factor(levels = c("Asia", 
             "Africa", "Oceania", "Europe", 
             "North America", "South America")), 
         `2022 Population` = col_number(), 
         `2020 Population` = col_number(), 
         `2015 Population` = col_number(), 
         `2010 Population` = col_number(), 
         `2000 Population` = col_number(), 
         `1990 Population` = col_number(), 
         `1980 Population` = col_number(), 
         `1970 Population` = col_number(), 
         `Area (km²)` = col_number(), `Density (per km²)` = col_number(), 
         `Growth Rate` = col_number(), `World Population Percentage` = col_number()))
```

We just changed the data types for each population year to *numeric*. Now we can run our actual analysis.

### Largest and smallest countries in the world

```{r}
large <- world_population %>%  arrange(desc(`2022 Population`))
head(large)
```

The Six Largest countries in the world are China, India, US, Indonesia, Pakistan, and Nigeria.

```{r}
small <- world_population %>% arrange(`2022 Population`)
head(small)
```

The Six smallest countries in the world are Vatican City, Tokelau, Niue, Falkland Islands, Monsterrat, and Saint Pierre and Miquelon.

### Percentage increase in Population from 1970 to 2022

```{r}
new_column <- mutate(world_population, percent_change = round(((`2022 Population` - `1970 Population`)/ `1970 Population`) * 100))
per_increase <- new_column %>% select(`Country/Territory`, Continent, `2022 Population`, `1970 Population`, percent_change) %>% 
  arrange(desc(percent_change))
head(per_increase)
```

The United Arab Emirates experienced the largest population increase from 1970 to 2022 rising over 3000%. The reasons for this are not examined in this analysis. 5 out of the top 6 countries with the highest percentage increase are located in Africa or Asia.

### Countries with least increase in population during this same period.

```{r}
new_column %>% select(`Country/Territory`, Continent, `2022 Population`, `1970 Population`, percent_change) %>% arrange(percent_change)
```
Some countries like the Vatican, Latvia and Georgia saw a decrease in their populations.


The total world population is:

```{r}
sum(world_population$`2022 Population`)
```

We can graph it in a bar chart by continent

```{r}
ggplot(data = sum_cont) + geom_col(mapping = aes(x = Continent, y = sum_2022, fill = Continent)) +
  ggtitle('Total World Population by Continent') + labs(x = 'Continent', y = 'Population')
```

### \# Population by continent

```{r}
total_pop <- world_population %>% group_by(Continent) %>% summarise(
  sum_2022 = sum(`2022 Population`),
  sum_2020 = sum(`2020 Population`),
  sum_2015 = sum(`2015 Population`),
  sum_2010 = sum(`2010 Population`),
  sum_2000 = sum(`2000 Population`),
  sum_1990 = sum(`1990 Population`),
  sum_1980 = sum(`1980 Population`),
  sum_1970 = sum(`1970 Population`)
)
View(total_pop)
```

### Continent by growth rate from 1970-2022

```{r}
cont_growth <- total_pop %>% mutate(percent_change = ((sum_2022 - sum_1970)/sum_1970) * 100) %>% arrange(desc(percent_change))
head(cont_growth)
```

Africa had the most population growth of all continents. The Least growth was experienced by Europe whose population rose by only 13.3% through 1970 to 2022.
