---
title: 'NYC Weather Analysis'
author: 'Charles Carnegie'
date: '27 July 2017'
output:
 html_document:
   number_sections: true
   toc: true
   fig_width: 7
   fig_height: 4.5
   theme: cosmo
   highlight: tango
   code_folding: hide
---
  
```{r}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, error=FALSE)
#________________________________________________________________________
# Note:   This script is written to be rendered into Rmd and HTML
#         The #' comments is stripped off when the code is rendered
#         Different types of blocks are market with lines
#________________________________________________________________________
```

# Introduction

The purpose of this kernel is to investigate the impact of weather on trip duration.
Heads or Tails has produced a more definitive analysis of the problem @ https://www.kaggle.com/headsortails/nyc-taxi-eda-update-the-fast-the-classified

As a new contributor, my goal is to build on his analysis by more completely analysing the impact of weather
To date, I have only been able to improve model fit by less than 1%
Questions, suggestions and comments would be welcome.

## Approach

The approach I have taken is to merge all the rides into one-hour bins and then match them to the METARS weather set.

The logic is:
 
1. Demonstrate that rides vary significantly by hour of the week. I.e. There are 24 x 7 or 168 separate hours of the week with different charecteristics.  

2. Then look at the impact of weather on ride features.

3. Then to calculate how much of ride variation can be explained by weather conditions, hour by hour.

4. Finally to figure out how to extract the key weather variables so that they can be used by other teams in the general model.

## Acknowledgements

Thanks to Heads or Tails @ https://www.kaggle.com/headsortails from his great general intro kernel and graphics

Thanks to  CarlesBalsach @ https://www.kaggle.com/cabaki for the weather dataset.

Thanks to WTI200 @ https://www.kaggle.com/wti200 for his general kernal.

# Setup



```{r }
#________________________________________________________________________

library(tidyverse)
library(lubridate)
library(geosphere)
library(dplyr)
library(ggplot2)
library(scales)
library(grid)
library(forcats)

#________________________________________________________________________
#
# Read with Kaggle input addresses
#
KNYC_Metars <- read_csv('../input/knycmetars2016/KNYC_Metars.csv')
train <- read_csv('../input/nyc-taxi-trip-duration/train.csv')
#test <- read_csv('../input/nyc-taxi-trip-duration/test.csv')


#________________________________________________________________________
#
# Add in the multiplot function
#
multiplot <- function(..., plotlist = NULL, file, cols = 1, layout = NULL) {
  require(grid)
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  if (numPlots == 1) {
    print(plots[[1]])
  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
quad_layout <- matrix(c(1,2,3,4),2,2,byrow=TRUE)
#________________________________________________________________________
```


# Join the Weather Data

Join the weather data and trip data together based on the mid trip time

```{r }
#________________________________________________________________________
# 
# Merge the training data with KNYC_Metars weather data
# (based on the mid point of the trip)
#   
train$datetime <- (train$pickup_datetime + 
                         ((train$dropoff_datetime - 
                             train$pickup_datetime) / 2)) 
train$datehour <- as_datetime(cut(train$datetime,breaks = "hour"))

#
KNYC_Metars$datehour <- as.POSIXct(KNYC_Metars$Time,"%d/%m/%Y %H:%M",tz="UTC")
#
weatherdata <- left_join(train, 
                          KNYC_Metars, 
                          by.x = train$datehour,
                          by.y=KNYC_Metars$datehour)

# Fill in unknown weather conditions
weatherdata[is.na(weatherdata$Conditions),"Conditions"] <- "Unknown" 
weatherdata[is.na(weatherdata$Events),"Events"] <- "Unknown"
#________________________________________________________________________
```


# Feature Engineering

The following new features are created in this code block:
- Trip_to_airport trips are identified
- hour, day, Hour of Week (how) - Assume that all hours are different
- distance in metres
- speed in km/hr
- blizzard - notes the 2016 blizzard which was atypical
- long (> )and short trips (less than 1 metre) are flagged

```{r }
#________________________________________________________________________
#
# Calculate time factors like Hour of the Week (how) and set as factors
#
weatherdata$day <- wday(weatherdata$datehour)
weatherdata$date <- date(weatherdata$datehour)
weatherdata$hour <- hour(weatherdata$datehour)
weatherdata$how <- as.factor(100*weatherdata$day+weatherdata$hour) 
#
weatherdata <- weatherdata %>%
  mutate(day = wday(datetime, label = TRUE),
         day = fct_relevel(day, c("Mon", "Tues", "Wed", "Thurs", "Fri", "Sat", "Sun")),
         weekday = (day %in% c("Mon","Tues","Wed","Thurs","Fri")),
         worktime = (hour %in% seq(8,18)) & (day %in% c("Mon","Tues","Wed","Thurs","Fri"))
  )
weatherdata$hour <- as.factor(weatherdata$hour)
#
# Calculate trip distance speed and bearing
#
pick_coord <- weatherdata %>%
  select(pickup_longitude, pickup_latitude)
drop_coord <- weatherdata %>%
  select(dropoff_longitude, dropoff_latitude)

weatherdata$trip_dist <- distCosine(pick_coord, drop_coord)
weatherdata$bearing  <- bearing(pick_coord, drop_coord)
weatherdata$speed  <- weatherdata$trip_dist/weatherdata$trip_duration*3.6
#
# Flag trips to the airport
#
jfk_coord <- tibble(lon = -73.778889, lat = 40.639722)
la_guardia_coord <- tibble(lon = -73.872611, lat = 40.77725)
#
weatherdata$jfk_dist_pick <- distCosine(pick_coord, jfk_coord)
weatherdata$jfk_dist_drop <- distCosine(drop_coord, jfk_coord)
weatherdata$lg_dist_pick <- distCosine(pick_coord, la_guardia_coord)
weatherdata$lg_dist_drop <- distCosine(drop_coord, la_guardia_coord)
#
# Trips with a dropoff or pickup within 2km of the airport
weatherdata <- weatherdata %>%
  mutate(jfk_trip = (jfk_dist_pick < 2e3) | (jfk_dist_drop < 2e3),
         lg_trip = (lg_dist_pick < 2e3) | (lg_dist_drop < 2e3)
  )
#
# Label the 2016 blizzard
#
weatherdata <- weatherdata %>%
  mutate(blizzard = !(date < ymd("2016-01-22") | (date > ymd("2016-01-29")) ))
#
# Set the weather conditions as factors
#
weatherdata$weather <- as.factor(weatherdata$Conditions)
weatherdata$events <- as.factor(weatherdata$Events)
#
# Reorder the levels: from best to worst weather
#
# The weather factor levels are reordered as follows:
# [1] "Clear"               
#[10] "Partly Cloudy"
# [8] "Mostly Cloudy"      
# [9] "Overcast"
# [2] "Haze"
#[12] "Scattered Clouds"   
#[14] "Unknown" 
# [6] "Light Rain"          
# [7] "Light Snow"
#[11] "Rain"
# [5] "Light Freezing Rain"
#[13] "Snow"
# [3] "Heavy Rain"
# [4] "Heavy Snow"         

weatherdata$weather = factor(weatherdata$weather,
                              levels(weatherdata$weather)[c(1,10,8,9,2,12,14,6,7,11,5,13,3,4)])
#
#________________________________________________________________________
```


# Data Cleansing


```{r }
#________________________________________________________________________
# 
# This next piece of code is from Heads or Tails and reflects his analysis 
# it takes out about 8,000 records
#
weatherdata <- weatherdata %>%
  filter(trip_duration < 22*3600,
         trip_dist > 0 | (near(trip_dist, 0) & trip_duration < 60),
         jfk_dist_pick < 3e5 & jfk_dist_drop < 3e5,
         trip_duration > 10,
         speed < 100)

#________________________________________________________________________
```


# Visual Analysis

## Plot ride data by hour of the week
The data below shows that the number, speed and distance of rides varies through the day and between week and weekend days.


```{r }
#________________________________________________________________________
# 
# Sumarise the data by day and hour
#
trip_avgs <- weatherdata%>%
  group_by(day, hour,weekday)%>%
  summarise(no_trips=n(),
            avg_trip_dist=mean(trip_dist),
            avg_trip_dur = mean(trip_duration),
            trip_dist_var = sd(trip_dist))
#
trip_avgs$trip_t  <- trip_avgs$avg_trip_dist/trip_avgs$trip_dist_var
#
#  Plot the data 
#
g1 <- trip_avgs %>%
  group_by(hour) %>%
  ggplot(aes(hour,avg_trip_dur)) +
  geom_point(aes(color = weekday))+
  geom_smooth(aes())+
  labs(x = "Average daily results by hour of the day", y = "Average Trip Duration")

g2 <- trip_avgs %>%
  group_by(hour, weekday) %>%
  ggplot(aes(hour,avg_trip_dist)) +
  geom_point(aes(color = weekday))+
  geom_smooth(aes())+
  labs(x = "Average daily results by hour of the day", y = "Average Trip Distance")

g3 <- trip_avgs %>%
  group_by(hour, weekday) %>%
  ggplot(aes(hour,no_trips)) +
  geom_point(aes(color = weekday))+
  geom_smooth(aes())+
  labs(x = "Average daily results by hour of the day", y = "Number of rides started")

g4 <- trip_avgs %>%
  group_by(hour, weekday) %>%
  ggplot(aes(hour,trip_t)) +
  geom_point(aes(color = weekday))+
  geom_smooth(aes())+
  labs(x = "Average daily results by hour of the day", y = "Variation in trip distance (mean/sd)")

g1
g2
g3
g4
multiplot(g1,g2,g3,g4,layout=quad_layout)
#________________________________________________________________________
```


## Summarise and plot the weather conditions


```{r }
#________________________________________________________________________
#   
weather_days <- weatherdata%>%
  group_by(date,day,hour,how,weather,Precip,Visibility,events)%>%
  summarise(num_trips=n(),
            tot_trip_dur = sum(trip_duration),
            tot_trip_dist = sum(trip_dist))
#  
g5 <- weather_days %>%
  group_by(weather) %>%
  count() %>%
  ggplot(aes(weather, n, fill = weather)) +
  geom_col() +
  scale_y_sqrt()+
  labs(x = "Weather conditions in NYC 2016/17", y = "Occurance in Hours ")

g6 <- weather_days %>%
  group_by(events) %>%
  count() %>%
  ggplot(aes(events, n, fill = events)) +
  geom_col() +
  scale_y_sqrt()+
  labs(x = "Weather events in NYC 2016/17", y = "Occurance in Hours ")

g5
g6
#________________________________________________________________________
```


## Summarise and plot the impact of weather conditions on rides


```{r }
#_______________________________________________________________________________
#
# Plot the average weather by hour and the impact of weather
#
weather_avgs <- weather_days%>%
  group_by(how,day,hour,weather)%>%
  summarise(no_of_records=n(),
            tot_num_of_trips=sum(num_trips),
            tot_trip_dur = sum(tot_trip_dur),
            tot_trip_dist = sum(tot_trip_dist))
#
# Calculate the averages for each hour and condition
#
weather_avgs$trips_per_hour <- weather_avgs$tot_num_of_trips/weather_avgs$no_of_records
weather_avgs$avg_trip_dist <- weather_avgs$tot_trip_dist/weather_avgs$tot_num_of_trips
weather_avgs$avg_trip_dur <- weather_avgs$tot_trip_dur/weather_avgs$tot_num_of_trips
weather_avgs$avg_trip_speed <- weather_avgs$tot_trip_dur/weather_avgs$tot_trip_dist

#
b1 <- weather_avgs %>%
  ggplot(aes(weather, trips_per_hour, color = weather)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(y = "Trips/Hour", x = "Weather Conditions")

b2 <- weather_avgs %>%
  ggplot(aes(weather, avg_trip_dist, color = weather)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(y = "Average Trip Distance", x = "Weather Conditions")

b3 <- weather_avgs %>%
  ggplot(aes(weather, avg_trip_dur, color = weather)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(y = "Average Trip Duration", x = "Weather Conditions")

b4 <- weather_avgs %>%
  ggplot(aes(weather, avg_trip_speed, color = weather)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(y = "Average Trip Speed", x = "Weather Conditions")

b1
b2
b3
b4

multiplot(b1,b2,b3,b4,layout=quad_layout)
#________________________________________________________________________
```


# Modelling and Analysis


```{r }

#________________________________________________________________________
```


# Summary

Including weather data improves the model as follows:


```{r }
#________________________________________________________________________
#
# The following values are calculated on the home machine as the model exceeds the Kaggle memory requirement
base.r.squared <- 0.0698654
weather.r.squared <- 0.07020437

print(paste("R Squared for base model =",base.r.squared))
print(paste("R Squared for weather model =",weather.r.squared))
improvement <- weather.r.squared-base.r.squared
print(paste("Weather features improve model by",percent(improvement)))

#________________________________________________________________________
```

